# Tidyverse in packages

* How to refer to functions (recap)
* Tidy evaluation
  * Data masking
  * Tidy select
* Data in packages

Draws heavily on <https://ggplot2.tidyverse.org/dev/articles/ggplot2-in-packages.html>,
(thanks to [Dewey Dunnington](http://github.com/paleolimbot)) and <https://dplyr.tidyverse.org/dev/articles/programming.html>.

I'm going to focus on ggplot2, with a dash of dplyr, 
but the same principles apply to basically any function in the tidyverse.

## Recap 

**Your turn**: Turn this code in to package. 
Use what you learned yesterday to get as much as possible of R CMD check to work. 
Copy and paste from github!

```{r}
library(ggplot2)
library(dplyr)

# Create a local copy of first 10 rows of mtcars data frame
# We'll come back how you should store this later
mtcars <- data.frame(
  mpg = c(21, 21, 22.8, 21.4, 18.7, 18.1, 14.3, 24.4, 22.8, 19.2),
  cyl = c(6, 6, 4, 6, 8, 6, 8, 4, 4, 6),
  disp = c(160, 160, 108, 258, 360, 225, 360, 146.7, 140.8, 167.6),
  hp = c(110, 110, 93, 110, 175, 105, 245, 62, 95, 123),
  drat = c(3.9, 3.9, 3.85, 3.08, 3.15, 2.76, 3.21, 3.69, 3.92, 3.92),
  wt = c(2.62, 2.875, 2.32, 3.215, 3.44, 3.46, 3.57, 3.19, 3.15, 3.44),
  qsec = c(16.46, 17.02, 18.61, 19.44, 17.02, 20.22, 15.84, 20, 22.9, 18.3),
  vs = c(0, 0, 1, 1, 0, 1, 0, 1, 1, 1),
  am = c(1, 1, 1, 0, 0, 0, 0, 0, 0, 0),
  gear = c(4, 4, 4, 3, 3, 3, 3, 4, 4, 4),
  carb = c(4, 4, 1, 1, 2, 1, 4, 2, 2, 4)
)

# cyl_plot()
ggplot(mtcars) + 
  geom_bar(aes(cyl)) + 
  coord_flip()

# cyl_sum()
mtcars %>% 
  group_by(cyl) %>% 
  summarise(n = n(), mpg = mean(mpg))
```

**Hints:**

* `create_package("~/desktop/tidytest")`
* `use_mit_license()` / `use_gpl3_license()`
* make two functions: `cyl_plot()` and `cyl_summary()`
* make it work - remember dependency stuff (`use_something()`, and `::`)
* get check passing as cleanly as you can
* GREEN post it when done; PINK post it for problems

**Your goal:**  `0 errors ✓ | 0 warnings ✓ | 1 note x`:
```
> checking R code for possible problems ... NOTE
  cyl_plot: no visible binding for global variable ‘cyl’
  cyl_sum: no visible binding for global variable ‘cyl’
  cyl_sum: no visible binding for global variable ‘mpg’
  Undefined global functions or variables:
    cyl mpg

0 errors ✓ | 0 warnings ✓ | 1 note x
```

<!-- Commit up to here before starting -->

---

To get it working:

* Remove calls to `library()`, and instead call `use_package()` once
* Use `::` or `@importFrom` to access the functions
* `@importFrom magrittr %>%` to access the pipe

Using magrittr increases the explicit dependencies, 
but doesn't change the implicit dependencies.

---

<!-- Commit tidytest -->

## Data masking

Data masking is one kind of tidy evaluation - 
the framework that tidyverse packages use to make **doing** data analysis easier.

**Your Turn**: The following function calls use variables 
`mtcars`, `cyl`, `mpg`, `starwars`, `homeworld`, and `species`.
There's really two types of variables here. What are they? What's the difference?
```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(n = n(), mpg = mean(mpg))

starwars %>% 
  filter(homeworld == "Naboo", species == "Human")
```

---

* `mtcars`, `starwars` are "programming" variables 
    that exist in the environment, **env-variables**.  

*  `cyl`, `mpg`, `homeworld`, and `species` are "statistical" variables 
    that exist inside data, **data-variables**.

At its heart tidyeval blurs the line between 
variables in an environment (env-variables), and 
variables in a data frame (data-variables). 

### (+) Data-masking makes code more concise

This blurring gives much more concise code, 
e.g. compare to the base equivalent: 
```{r}
starwars[starwars$homeworld == "Naboo" & starwars$species == "Human", ,]
```
Here we have to be explicit about `homeworld` and `species` being data-variables 
by prefacing with `starwars$`.

### (-) Data-masking makes some things harder

Two problems:

1. `R CMD check` can't find data-variables in the environment, so it complains
2. It's hard to use indirection; i.e. storing a data-variable in an env-variable

Tidy evaluation is implemented by the rlang package, 
primarily through `enquo()` and `eval_tidy()`.  

### 1. Fix with rlang's `.data` pronoun

To fix the `R CMD check` note, 
we need to be more specific that `cyl` and `mpg` are data-variables.

rlang provides a `.data` pronoun.
Use it in a function that uses data masking 
to be explicit about a data variable.

Two steps to use in a package:

* Need to import `.data` from `rlang`, and include rlang in Imports
* Refer to data variables as `.data$var`

```{r}
#' @importFrom rlang .data
cyl_plot <- function(){
  ggplot2::ggplot(mtcars) +
    ggplot2::geom_bar(ggplot2::aes(.data$cyl)) +
    ggplot2::coord_flip()
}
```

**Your turn**: Add this function and get R CMD check passing:

```{r}
#' @export
#' @importFrom rlang .data
diamonds_summary2 <- function(bins) {
  ggplot2::diamonds %>% 
    ggplot2::ggplot(ggplot2::aes(carat, price)) + 
    ggplot2::geom_hex(bins = bins)
}
```

### 2. It's hard to use indirection

Can we create `bar_plot()` that works like `cyl_plot()`,
except instead of always plotting the `cyl` column,
the user provides the variable to be plotted?

E.g. `bar_plot(cyl)` should give the same result as `cyl_plot()`

Try:
```{r}
bar_plot <- function(var){
  mtcars +
    ggplot2::geom_bar(ggplot2::aes(.data$var)) +
    ggplot2::coord_flip()
}
```
``Error: Column `var` not found in `.data` ``

### 2. Fix by making user provide character string and use `.data`

Make user provide a string `bar_plot("cyl")`.
We can use `.data` but now provide `var` in `[[`:
```{r}
bar_plot <- function(var){
  mtcars +
    ggplot2::geom_bar(ggplot2::aes(.data[[var]])) +
    ggplot2::coord_flip()
}
```

But what if you want the function to work like tidyverse functions: `bar_plot(cyl)`?

### 2. To work like tidyverse function, fix with embracing `{{`

Need to tell `aes()` that env-variable `var` contains a data-variable: 
do that with embracing, `{{`.

```{r}
bar_plot <- function(var){
  ggplot2::ggplot(mtcars) +
    ggplot2::geom_bar(ggplot2::aes({{ var }})) +
    ggplot2::coord_flip()
}
```

Inspired by glue
```{r}
name <- "Charlotte"
glue::glue('Hello, my name is { name }')
```


Other alternatives: make user provide `aes()` call, use `...`.

**Your turn**: Make these functions work and pass R CMD check

```{r}
cyl_sum <- function(x){
  mtcars %>%
  dplyr::group_by(.data$cyl) %>%
  dplyr::summarise(n = dplyr::n(), avg = mean(x))
}
cyl_sum(wt)
```

```{r}
hexbin <- function(df, x, y, bins = 10) {
  df %>% 
    ggplot2::ggplot(ggplot2::aes(x, y)) + 
    ggplot2::geom_hex(bins = bins)
}
hexbin(diamonds, carat, price)
```

```{r}
group_summary <- function(df, grp, x) {
  df %>% 
    group_by(grp) %>% 
    summarise(
      n = n(),
      min = min(x, na.rm = TRUE),
      max = max(x, na.rm = TRUE)
    )
}
group_summary(mtcars, cyl, mpg)
```

**Challenge**: Allow `group_summary()` to take two grouping variables. 

---

```{r}
cyl_sum <- function(x){
  mtcars %>%
  dplyr::group_by(.data$cyl) %>%
  dplyr::summarise(n = dplyr::n(), avg = mean({{ x }}))
}
cyl_sum(wt)
```

```{r}
hexbin <- function(df, x, y, bins = 10) {
  df %>% 
    ggplot2::ggplot(ggplot2::aes({{ x }}, {{ y }})) + 
    ggplot2::geom_hex(bins = bins)
}
hexbin(diamonds, carat, price)
```

```{r}
group_summary <- function(df, grp, x) {
  df %>% 
    group_by({{ grp }}) %>% 
    summarise(
      n = n(),
      min = min({{ x }}, na.rm = TRUE),
      max = max({{ x }}, na.rm = TRUE)
    )
}
group_summary(mtcars, cyl, mpg)
```
```{r}
group_summary <- function(df, grp1, grp2, x) {
  df %>% 
    group_by({{ grp1 }}, {{ grp2 }}) %>% 
    summarise(
      n = n(),
      min = min({{ x }}, na.rm = TRUE),
      max = max({{ x }}, na.rm = TRUE)
    )
}
group_summary(mtcars, cyl, gear, mpg)
```

```{r}
group_summary <- function(df, x, ...) {
  df %>% 
    group_by(...) %>% 
    summarise(
      n = n(),
      min = min({{ x }}, na.rm = TRUE),
      max = max({{ x }}, na.rm = TRUE)
    )
}
group_summary(mtcars, mpg, cyl, gear)
group_summary(mtcars, mpg, cyl)
```
---

### Data Masking - Summary

dplyr: `arrange()`, `filter()`, `group_by()`, `mutate()`, and `summarise()`  
ggplot2’s: `aes()`

In future, arguments that use data masking 
will be indicated in docs with `<data-masking>`.

* You (as developer) know the data-variable is `x`: `arrange(df, .data$x)`
* Name of data-variable is a **character string** in an env-variable `x`: 
`arrange(df, .data[[x]])`
* Data-variable is passed as a function argument `x`: `arrange(df, {{ x }})`

Remember `...` can be passed through - no special treatment required.

## Tidy Selection 

The other half of tidy evaluation, used in functions like:
`select()`, `rename()`, `pivot_longer()` ...

Implemented with tidyselect package: 
choose variables based on their position, name, or type.

Run into similar problem with indirection;
```{r}
my_select <- function(df, x) {
  df %>% 
    select(x)
}
my_select(mtcars, cyl)
```
<!-- Example fixed with embracing -->
```{r}
my_select <- function(df, x) {
  df %>% 
    select({{ x }})
}
my_select(mtcars, cyl)
```

```r
vars <- c("cyl", "gear")
mtcars %>% 
  select(one_of(vars))
```

### Tidy select - Summary

Some functions have "select" semantics - 
you don't give them operations that work with values, 
but operations that work with column names:

* You (as developer) know the column is `x`:  `select(df, one_of("x"))`
* Name of column is a **character string** in an env-variable `x`: 
`select(df, one_of(x))`
* Column is passed as a function argument `x`: `select(df, {{ x }})`

Remember `...` can be passed through - no special treatment required.

https://dplyr.tidyverse.org/dev/articles/programming.html

Undergoing some changes for dplyr 1.0.0:

* `one_of()` replaced by `any_of()` and `all_of()` - 
what should happen if variable doesn't exist in data frame?

* Tidy selection inside data masking functions? New function `across()`, used with `mutate()` and `summarise()`,  replaces `mutate_at()`, `summarise_at()`, `mutate_if()` etc.

## Data

To include data in a package:

* `use_data_raw("name_of_dataset")`
* creates an R file in `data-raw` with same name as your dataset
* `usethis::use_data()` to create the binary data file that R uses (wrapper around `save()`)

* To document: 
  * `use_r("data")`
  *
      ```R
      #' Title
      #'
      #' @description
      #' Data dictionary:
      #'
      #' * `var1`: description1
      "name_of_dataset"
      ```

A little more information at <https://roxygen2.r-lib.org/articles/rd.html#datasets>
