# Function interface

Goal for today is focus on the interface of functions, because I think this is an important change in mindset when you start developing functions. You have to think how they fit together into a greater whole.

* interface = outside
* implementation = inside

More (very rough thoughts at) <http://design.tidyverse.org/>

**Your turn**: What are the properties of a function that define its interface? (There are at least 9!)

```{r}
countdown::countdown_fullscreen(3)
```

* Name
* Inputs (i.e. argument)
  * Names
  * Types
  * Defaults
  * Evaluation (NSE, tidy, ...)
* Output
* Errors (and other conditions)
* Side-effects
  * Saving files
  * Messaging to user

```{r setup}
library(stringr)
```

## base R vs stringr

NB: I don't think string functions in base R are bad or wrong. But they grew organically over multiple years, and I think we can do better with some systematic anaylsis. Similarly, stringr functions aren't perfect, and there are definitely mistakes. Really important to critique existing work in order to get better. (And base string function have a good collection of problems)

> "Each [function] is perfect the way it is .... and it can use a little improvement."
> --- Shunryu Suzuki

**Your turn**: What are some of the challenges of the base string functions? Think particularly about names, arguments, and outputs.

```{r, eval = FALSE}
strsplit(x, split, ...)
grep(pattern, x, value = FALSE, ...)
grepl(pattern, x, ...)
sub(pattern, replacement, x, ...)
gsub(pattern, replacement, x, ...)
regexpr(pattern, text, ...)
gregexpr(pattern, text, ...)
regexec(pattern, text, ...)
substr(x, start, stop)
nchar(x, type, ...)
```

* Names
  * No common theme. 
  * No common prefix. 
  * Names are concise at expense of expressiveness. (who is Greg?)
  
* Arguments
  * Argument names & order are not consistent
  * Data isn’t the first argument (2nd or 3rd). 
  * Sometimes `text`, sometimes `x`. 
  * Not always independent.

* Outputs:
  * Can’t feed output of `gregexpr()` into `substr()`

* Errors?
* Side-effects?

## Names

> "A rose by any other name would smell as sweet."
> --- Shakespeare

> "A function by any other name would not smell as sweet."
> --- me

Name of function is really important part of its interface because it's the mental hook upon which the functionality is hang. Me and my team spend a lot of time dicussing sames.

### Function should be verbs (usually)

Because they _do_ something

**Your turn**: Match the following functions and arguments to their stringr equivalents. What are some issues with the base functions? What are some issues with the stringr equivalents?


```{r}
grepl(value = TRUE) # -> str_subset()
grep()              # -> str_which() 
grepl()             # -> str_detect()
sub()               # -> str_replace()
gsub()              # -> str_replace_all()
strsplit()          # -> str_split()
regexpr()           # -> str_locate()
```

Some problems: what's the difference between `str_detect()`, `str_which()`, and `str_subset()`? (Similarly problem with `dplyr::transmute()` and `dplyr::mutate()`) Differ primarily in their output type. English verbs too similar:

```{r}
str_detect() # str_something_lgl() ?
str_which()  # str_something_int() ?
str_subset() # str_something_chr() ?
```

Another way to think about it is that some function work with chararacters in a string, which others work with strings in a character vector.

```{r}
str_locate_vec()
str_locate_str()
```

---

**Your turn**: Most ggplot2 functions aren't verbs. Is this a good idea? If they were to be verb, what would you call them? Discuss

---

Verbs aren't always essential; nouns can work when you're building up objects

```{r, eval = FALSE}
library(ggplot2)

geom_line()
scale_x_continuous()
coord_fixed()
```

Or extracting properties:

```{r, eval = FALSE}
str_length()
```

And sometimes there just isn't a good verb:

```{r}
str_to_lower()
str_to_upper()
```

### Prefixes and suffixes

* `snake_case` vs `camelCase`. I don't think one is intrinsically better  
  than the other. But consistency is VERY IMPORTANT. I use snake_case
  (mostly because of my background with ruby). Please don't be like
  `install.packages()`

* Use prefixes to group together related functions. 
  Main advantage is autocomplete. 
  Shiny is a counter example.
  Not necessary to have common prefix for every function in package.

* Use suffixes for variations on a theme.
    * `str_replace()` and `str_replace_all()`
    * `str_split()` and `str_split_fixed()`

**Your turn**: What stringr (and other tidyverse) functions don't use these principles?

---

```{r, eval = FALSE}
# Don't start with str_
invert_match()
word()
fixed()
regexp()
```

---

When should you make multiple functions rather than using an argument? If it changes the _type_ of output, always make it a function. If the argument would interact with other arguments, probably should be a function.  

If there are many options, an argument probably better. Sometimes only obvious in retrospect.

```{r}
# Sarted with
str_to_upper()
str_to_lower()
# Then added
str_to_title()
str_to_sentence()

# probably should be
str_capitalise(x, to = "upper")
str_capitalise(x, to = "lower")
str_capitalise(x, to = "title")
str_capitalise(x, to = "sentence")
# BUT
str_capitalize() # ?
```

## Arguments

* Types
* Required arguments
* Hidden arguments
* Enumms
* Keeping arguments independent

### Types

**Your turn**: Can divide the following arguments into function groups? To get started, I think `x` and `text` form one group.

```{r}
strsplit(x, split, fixed = TRUE)
grepl(pattern, x, fixed = TRUE, ignore.case = TRUE)
gsub(pattern, replacement, x, fixed = TRUE, ignore.case = TRUE)
gregexpr(pattern, text, fixed = TRUE)
substr(x, start, stop)
```

(These are not very good but they're the best we have)

* Data: `x` or `text`. Primary vector of data. Often called `x` or `data`.
  Required. Controls shape/type of output

* Descriptors: `pattern`, `replacement`, `start`, `stop`. 
  Required. Usually only 2 or 3 more
  V. important.

* Details: `fixed`, `ignore.case`, `perl`, `useBytes` etc
  Optional.
  Often a scalar.

---

These categories are important because they guide how we call a function. And how we organise

* Data
* Descriptors
* Dots (if used)
* Details

Benefits:

* Works with pipe.
* Ordered from most important to least important
* Required arguments come before optional. 
* Forces details to be named.

### Required

**Your turn**: Which arguments are required and optional in these functions? You'll need to look at signature and read docs.

```{r}
lm()
sample()
diag()
str_replace_all()
```

---

Stick to this simple rule:
* Required <-> no default
* Optional <-> has default

Warning sign: use of `missing()`. If know what it is, don't use it. If you don't; don't learn it. 

(There are occassional exceptions where you have two mutually exclusive arguments, e.g. `rvest::html_node(css, xpath)`)

### Hidden

**Your turn**: Do these functions always return the same results?

```{r}
data.frame(x = "abc")
tolower("I")
sort(letters)
```

---

```{r}
str_to_lower("I")
str_to_lower("I", locale = "tr")

# https://en.wikipedia.org/wiki/Swedish_alphabet
str_sort(c("a", "o", "z", "å", "ø"))
str_sort(c("a", "o", "z", "å", "ø"), locale = "se")
```

<https://www.vice.com/en_us/article/zmjwda/a-code-glitch-may-have-caused-errors-in-more-than-100-published-studies>

---

Obviously bad; generally either straightforward to avoid or you don't know about them. Mostly related to locale.

Americentrism vs code that works everywhere.

```{r}
readr::locale()
```


### Enums

**Your turn**: What do these functions have in common?

```{r}
trimws()
str_trunc()
rank()
```

---

This is a really useful technique if you have an argument that takes a small set of possible values. Key is use of `match.arg()`

### Independence

```{r}
grepl(".x", "XY.X", fixed = TRUE, ignore.case = TRUE)
grepl(".x", "XY.X", fixed = TRUE, perl = TRUE)
```

```{r, error = TRUE}
rgamma(5, shape = 1, rate = 2, scale = 1/2)
rgamma(5, shape = 1, rate = 2, scale = 2)
```

stringr takes a different approach:

```{r}
str_detect("XY", fixed("x"))
str_detect("XY", regex("x", multline = TRUE))
str_detect("XY", fixed("x", ignore_case = TRUE))
str_detect("X Y", boundary("word"))
```

### Type stability

Not a big problem with string functions.

**Your turn**: Assuming this code works (with out errors or warnings):

```{r}
y <- sum(x)
```

* What do you know about `x`?
  * Numeric or logical vector (or matrix/array)
  
* What do you know about `y`?
  * Length 1
  * Numeric

**Your turn**: Assuming that `df` is a data frame, and this code works (with out errors or warnings):

```{r}
df[, i]
```

* What do you know about `i`?
* What do you know about the result?

## Output

### Working with input

Exception is (I think) two mutually acceptable arguments, or when one function returns a complex object that could be pulled apart into multiple objects.

```{r}
str_sub(x, 1, 2)
str_sub(x, str_locate(x, "a+"))
```

```{r}
x <- c("bbaab", "bbb", "bbaaba")
loc <- regexpr("a+", x)
substr(x, loc, loc + attr(loc, "match.length") - 1)

# And only works because this returns ""
substr(x, -1, -3)

# regmatches() has a different problem
regmatches(x, loc)
```

### Invariants

Useful to be able to describe the output of a function using a few simple statements that we'll call invariants. For example take `str_sub()`.

* Output is always a character vector
* Output has length `max(length(string, start, end))` (using recycling rules)

What about `str_detect()`?

* What is the output type?
* What size is it?

**Your turn**: How can you describe the output of `grep()`?

---

As a general principle, I think functions with smaller/shorter/simpler/standard invariants are better because they're more likely to stay in your head. But more on that in a little bit. 

Often important to think about invariants when things are bit unusual. 

What happens when there isn't a match?

**Your turn**: Why can `str_locate()` return a matrix while `str_locale_all()` has to return a list? Why does it return a matrix and not a data frame? Why are the return values different when there is no match?

```{r}
x <- c(NA, "", "x", "y", "yy")
str_locate(x, "y")
str_locate_all(x, "y")
```

**Your turn**: Are these results consistent?

<!-- Expand on these - should analyse locate() and locate_all() inviarnats first -->

```{r}
str_extract(x, "y")
str_extract_all(x, "y")

str_split(x, "x")
str_split_fixed(x, "x", 2)
```

What would be consequences of changing this behaviour to be more consistent?