---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Dependencies 

!!! LIVE CODING BEGINS NOW !!!

<https://github.com/rstudio-conf-2020/build-tidy-tools>

## Motivation

```{r}
sd

x <- 1:10
sd(x)

var <- function(x) 100
sd(x)

my_sd <- function(x) sqrt(var(x))
my_sd(x)
```

### Warm ups

```{r}
x <- 1
f <- function() {
  y <- 2
  z <- 2
  g <- function() {
    z <- 3
    c(x, y, z)
  }
  y <- 4
  g()
}
f()
```

* What is an environment?
  * mapping from names to objects 
  * hierarchy - every environment has parent
  * data structure that powers scoping
  * can use environment as dictonaries/hashmaps

* How is an environment different from a list?
  * environments can't really have classes
  * environments have parents
  
```{r}
library(rlang)

e <- env()
e$a <- 1
e$a
e[[1]]

e$e <- e
e$e$e$e$e$e$e$a
lobstr::obj_addr(e)
e$f <- 10
lobstr::obj_addr(e)
str(e)

l <- list(a = 1)
lobstr::obj_addr(l)
l$l <- l
lobstr::obj_addr(l)
str(l)
```


* How can you see the contents of an environment?

```{r}
e
str(e)

ls(e)
names(e)
str(e$a)
str(as.list(e))

env_print(e)
```

```{r}
countdown::countdown_fullscreen(1)
```

## Scoping

```{r}
x

find_var <- function(name, env) {
  if (env_has(env, name)) {
    env_get(env, name)
  } else {
    find_var(name, env_parent(env))
  }
}

e1 <- env()
e1$a <- 11111
find_var("a", e1)
find_var("b", e1)
```

```{r}

find_var <- function(name, env) {
  if (identical(env, emptyenv())) {
    stop("object '", name, "' not found", call. = FALSE)
  } else if (env_has(env, name)) {
    # env[[name]] - equivalent
    env_get(env, name)
  } else {
    find_var(name, env_parent(env))
  }
}
find_var("a", e1)
find_var("b", e1)

```

```{r}
find_env <- function(name, env) {
  if (identical(env, emptyenv())) {
    stop("object '", name, "' not found", call. = FALSE)
  } else if (env_has(env, name)) {
    env
  } else {
    find_env(name, env_parent(env))
  }
}

find_env("a", e)
find_env("find_env", e)
find_env("sd", e)

```

```{r}
bench::mark(
  find_var("abort", globalenv()),
  find_var("is", globalenv()),
  check = FALSE
)
```

```{r}
sd
find_env("var", environment(sd))

my_sd
find_env("var", environment(my_sd))
```

