# Tidyverse in packages

* How to refer to functions (recap)
* Tidy evaluation
  * Data masking
  * Tidy select
* Data in packages

Draws heavily on <https://ggplot2.tidyverse.org/dev/articles/ggplot2-in-packages.html>,
(thanks to [Dewey Dunnington](http://github.com/paleolimbot)) and <https://dplyr.tidyverse.org/dev/articles/programming.html>.

I'm going to focus on ggplot2, with a dash of dplyr, 
but the same principles apply to basically any function in the tidyverse.

## Recap 

**Notes**: rstd.io/built-tt in `notes/4-tidyverse.Rmd` 

**Your turn**: Turn the code below in to a package. 
Use what you learned yesterday to get as much as 
possible of `check()` to work. 
Copy and paste from github! See **Hints** and **Goal** too.
Pink post-it for help, Green for reaching Goal.

```{r}
library(ggplot2)
library(dplyr)

# Create a local copy of first 10 rows of mtcars data frame
# We'll come back how you should store this later
mtcars <- data.frame(
  mpg = c(21, 21, 22.8, 21.4, 18.7, 18.1, 14.3, 24.4, 22.8, 19.2),
  cyl = c(6, 6, 4, 6, 8, 6, 8, 4, 4, 6),
  disp = c(160, 160, 108, 258, 360, 225, 360, 146.7, 140.8, 167.6),
  hp = c(110, 110, 93, 110, 175, 105, 245, 62, 95, 123),
  drat = c(3.9, 3.9, 3.85, 3.08, 3.15, 2.76, 3.21, 3.69, 3.92, 3.92),
  wt = c(2.62, 2.875, 2.32, 3.215, 3.44, 3.46, 3.57, 3.19, 3.15, 3.44),
  qsec = c(16.46, 17.02, 18.61, 19.44, 17.02, 20.22, 15.84, 20, 22.9, 18.3),
  vs = c(0, 0, 1, 1, 0, 1, 0, 1, 1, 1),
  am = c(1, 1, 1, 0, 0, 0, 0, 0, 0, 0),
  gear = c(4, 4, 4, 3, 3, 3, 3, 4, 4, 4),
  carb = c(4, 4, 1, 1, 2, 1, 4, 2, 2, 4)
)

# cyl_plot()
ggplot(mtcars) + 
  geom_bar(aes(cyl)) + 
  coord_flip()

# cyl_sum()
mtcars %>% 
  group_by(cyl) %>% 
  summarise(n = n(), mpg = mean(mpg))
```

**Hints:**

* `create_package("~/desktop/tidytest")`
* `use_mit_license()` / `use_gpl3_license()`
* make two functions: `cyl_plot()` and `cyl_summary()`
* make it work - remember dependency stuff (`use_something()`, and `::`)
* get check passing as cleanly as you can
* GREEN post it when done; PINK post it for problems

**Notes**: rstd.io/built-tt in `notes/4-tidyverse.Rmd` 

**Your goal:**  `0 errors ✓ | 0 warnings ✓ | 1 note x`:
```
> checking R code for possible problems ... NOTE
  cyl_plot: no visible binding for global variable ‘cyl’
  cyl_sum: no visible binding for global variable ‘cyl’
  cyl_sum: no visible binding for global variable ‘mpg’
  Undefined global functions or variables:
    cyl mpg

0 errors ✓ | 0 warnings ✓ | 1 note x
```
Extra task: web search skills to solve NOTE


**Your turn**

## Data Masking

* `mtcars`, `starwars` - datasets, variables available in the environment, environnment-varaibles, "programming" variables

* `cyl`, `mpg`,  `homeworld`, and `species` - columns in datasets, variables availble in data, data-variables, "statistical" variables

What are they are the two types of variables I'm referring to?
How are they different?

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(n = n(), mpg = mean(mpg))

starwars %>% 
  filter(homeworld == "Naboo", species == "Human")
```

## ++ Why does the tidyverse use data maskingn?

```{r}
starwars %>% 
  filter(homeworld == "Naboo", species == "Human")

starwars[starwars$homeworld == "Naboo" & starwars$species == "Human", ]
```

## -- Programming with the tidyverse is harder

* NOTE on check for data variables
* "indirection" is harder 

## 1. check() NOTE

If you refer to a data-variable inside your function,
preface with `.data`, rlang package

* Import `.data` from rlang, also in DESCRIPTION
* Refer to a data-varibale, `var`: `.data$var`

**Your Turn**:

Fix the package to get rid of NOTE about `cyl` and `mpg`
* Import `.data` from rlang, also in DESCRIPTION
* Refer to a data-varibale, `var`: `.data$var`

Add this to your package, and get it working (GREEN)
```{r}
#' @export
#' @importFrom rlang .data
diamonds_summary2 <- function(bins) {
  ggplot2::diamonds %>% 
    ggplot2::ggplot(ggplot2::aes(.data$carat, .data$price)) + 
    ggplot2::geom_hex(bins = bins)
}
```

## 2. Indirection is hard

Example of trying to use "indirection": I'm  indrectoyly tryinng to referring to the `gear` column.

One solution make user pass a string, use `[[`:
```{r}
bar_plot <- function(var){
  ggplot2::ggplot(mtcars) +
    ggplot2::geom_bar(ggplot2::aes(.data[[var]])) +
    ggplot2::coord_flip() +
    ggplot2::labs(x = var)
}
bar_plot("gear")
```

Better solution:  embracing, `{{ x }}` 
Look inside the environment variable x, 
and treat the contents like a data variable

```{r}
bar_plot <- function(var){
  ggplot2::ggplot(mtcars) +
    ggplot2::geom_bar(ggplot2::aes({{ var }})) +
    ggplot2::coord_flip() 
}
bar_plot(sqrt(gear))
```

## Your Turn:

**Your turn**: Make these functions work and pass R CMD check

```{r}
cyl_sum <- function(x){
  mtcars %>%
  dplyr::group_by(.data$cyl) %>%
  dplyr::summarise(n = dplyr::n(), avg = mean(x))
}
cyl_sum(wt)
```

```{r}
hexbin <- function(df, x, y, bins = 10) {
  df %>% 
    ggplot2::ggplot(ggplot2::aes(x, y)) + 
    ggplot2::geom_hex(bins = bins)
}
hexbin(diamonds, carat, price)
```

```{r}
group_summary <- function(df, grp, x) {
  df %>% 
    group_by(grp) %>% 
    summarise(
      n = n(),
      min = min(x, na.rm = TRUE),
      max = max(x, na.rm = TRUE)
    )
}
group_summary(mtcars, cyl, mpg)
```

**Challenge**: Allow `group_summary()` to take two 
grouping variables.

