---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Interface

* interface = outside of function
* implementation = inside of the function

<https://design.tidyverse.org>

**Your turn**: What are the properties of a funtion that define its interface? (There are at 9 properties)

* Name
* Arguments (inputs)
  * Names and order
  * Default values / required vs optional
  * Types
  * Evaluation (regular, NSE, tidy evaluation)
* Output
* Conditions
  * Errors
  * Warnings
  * Messages
* Side effect
  * print to console
  * making a plot
  * io - writing a file
  * change the random seed

## Regular expressions / string manipulation

stringr vs base R

```{r}
strsplit(x, split, ...)
grep(pattern, x, value = FALSE, ...)
grepl(pattern, x, ...)
sub(pattern, replacement, x, ...)
gsub(pattern, replacement, x, ...)
regexpr(pattern, text, ...)
gregexpr(pattern, text, ...)
regexec(pattern, text, ...)
substr(x, start, stop)
nchar(x, type, ...)
```

Challenges:
* Names
  * No common theme (g = global (mostly))
  * Short at the cost of expressiveness
  * Rely on familiarity with linux command line tools
  * No common prefix

* Inputs
  * Primary argument has inconsistent name: `text`, `x`
  * Most important argument doesn't come first (1st, 2nd, or 3rd)

* Output
  * `regexpr` returns a location, but you can't feed that location in `substr()`
  
## Names

## Function should be verbs (usually)

Match these base functions with their stringr equivalent: <https://stringr.tidyverse.org/articles/from-base.html>. What's good about the names? What's bad about them?

```{r}
grepl()            # -> str_detect()
grep()             # -> str_which()
grep(value = TRUE) # -> str_subset()
sub()              # -> str_replace()
gsub()             # -> str_replace_all()
strsplit()         # -> str_split()
regexpr()          # -> str_locate()
```

```{r}
str_detect() # -> str_something_lgl()
str_which()  # -> str_something_int()
str_subset() # -> str_something_chr()
```

```{r}
str_locate() # location of pattern within string
str_locate_pattern()
str_which()  # location of strings within vector
str_locate_vector()
```

**Your turn**: Look at ggplot2 - which functions are verbs? (`help(package = ggplot2)`) Is that good idea? If you were to make more functions into verbs, what verbs would you use?

```{r}
ggplot() %>%
  add_geom_point() %>%
  _scale_y_continous() %>%
  add_coord_equal()


```

```{r}
str_length()
str_to_lower()
str_to_title()
```

### Prefixes and suffixes

Guided by autocomplete:

* Use prefixes to group together related functions
* Use suffixes for variations on a theme

### Organising function name components

* Save dots only for S3 methods
* Otherwise be consistent! I like snake_case best; camelCase is fine.


```{r}
x <- structure(1:10, class = "test")
t(x)
# calls t.test()
```

### When to use a suffix vs. when to use an argument?

```{r}
str_replace_all()
# why not
str_replace(all = TRUE)

x <- c("aba", "aa","b")
str_locate(x, "a")
str_locate_all(x, "a")
# an argument should not generally change the
# type of the output

```

```{r}
str_to_lower()
str_to_upper()
str_to_title()
str_to_sentence()

str_capitalize(somthing = "lower")
str_capitalize(somthing = "upper")
str_capitalize(somthing = "title")
str_capitalize(somthing = "sentence")
```

Avoid UK vs US spelling differences

```{r}
ggplot2::scale_color_grey()

#' @export
#' @rdname summarise
summarize <- summarise
```

